Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2018-02-19 paper.tex

Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair being classified as a link or
non-link depending on their similarity. Linkage is usually performed in
a three-step process: first, groups of similar candidate records are
identified using indexing, then pairs within the same group are compared
in more detail, and finally classified. Even state-of-the-art indexing
techniques, such as Locality Sensitive Hashing, have potential
drawbacks. They may fail to group together some true matching records
with high similarity, or they may group records with low similarity,
leading to high computational overhead. We propose using \emph{metric
space indexing} (MSI) to perform \emph{complete} linkage, resulting in a
parameter-free process combining indexing, comparison and classification
into a single step delivering complete and efficient record linkage. An
evaluation on real-world data from several domains shows that linkage
using MSI can yield better quality than current indexing techniques,
with similar execution cost, without the need for domain knowledge or
trial and error to configure the process.

@@

To overcome data quality issues such as typographical errors and
variations (common in name and address values~\cite{Chr12}), approximate
string comparison functions (e.g. edit distance, the Jaro-Winkler
comparator, or Jaccard similarity~\cite{Chr12}) are used to compare
record pairs, leading to a vector of similarities (one similarity per
attribute compared) for each pair. These are used to classify the record
pairs into \emph{links} (where it is assumed both records correspond to
the same real-world entity) and \emph{non-links} (where they are assumed
to correspond to different entities). Various classification methods
have been employed in record linkage~\cite{Chr12,Don15}, ranging from
simple threshold-based to sophisticated clustering, supervised
classification techniques, and active learning approaches~\cite{Wan15}.

@@

Traditional linkage systems that perform indexing prior to comparison
and classification (on the left in \Cref{fig-rl-process}) add a further
complexity. Indexing, comparison and classification are often conducted
using algorithms and parameters selected using domain expertise,
followed by manual assessment of the linkage outcomes~\cite{Chr12}. If
the resulting link quality is too low for a certain application, the
process is repeated with different parameter settings or algorithms,
giving a time-consuming iterative process. The choice of an appropriate
indexing technique and parameters (including which attributes to use in
indexing) will significantly affect the final linkage outcome.

@@

Metric space indexing (MSI) is a complete technique with lower
computational cost than a brute force approach. It allows indexing,
comparison and classification to be combined into a single step (on the
right in \Cref{fig-rl-process}), making the process simpler, more
efficient and more effective than incomplete approaches.

The motivation for this work is
the Digitising Scotland project~\cite{Dibben2012}, which aims to
transcribe and link all civil registration events recorded in Scotland
between 1856 and 1973. This data set will include around 14 million
birth records, 11 million death records and 4 million marriage records.

@@

Locality sensitive hashing (LSH), proposed for efficient
nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
been used for record linkage indexing. Attribute values are hashed
multiple times, and blocks are created from those records that share
some hash values. \emph{HARRA}~\cite{Kim10} is a linkage approach based
on MinHash~\cite{Broder1997} and LSH which blocks, compares, and then
merges linked records iteratively. \cite{Steorts2014} evaluates two LSH
variations, concluding that to get good results, they must be tuned to
the datasets. This requires good ground truth data which may be
unavailable or expensive to obtain.

@@

\textbf{Brute force:} Every record in $\mathbf{D}_A$ is compared with
every record in $\mathbf{D}_B$. Each pair is classified as a link if the
distance between the records is less than or equal to the threshold.
This always finds all links, with complexity $ O(|\mathbf{D}_A| \times
|\mathbf{D}_B|)$.

\textbf{Traditional blocking:} The parameters are the set of blocking
keys and (optionally) the phonetic encodings applied to each attribute.
These are selected as described in~\cite{Chr12b}, exploiting knowledge
of the domain and of the data, and chosen with the intention of giving
the best possible results. Each record in $\mathbf{D}_A$ is placed into
the appropriate block based on its blocking key value. The algorithm
then iterates over the records in $\mathbf{D}_B$, and for each one,
compares it with each of the records from $\mathbf{D}_A$ in the block
with the same blocking key value.

\textbf{LSH-MinHash:} The parameters for LSH-Minhash
are~\cite{Broder1997} \emph{shingle size} ($l_{ss}$), \emph{band size}
($l_{bs}$) and \emph{number of bands} ($l_{nb}$). First, the attributes
of each record in $\mathbf{D}_A$ are concatenated, and the result
\emph{shingled} into a set of n-grams with $n = l_{ss}$. Next, a set of
deterministically generated hash functions are applied to each n-gram in
the set and the smallest result (the MinHash) of each hash application
is added to a signature for the record. The number of hashes used, and
thus the size of the signature, is set to $l_{nb} \times l_{bs}$.
Finally, the signature is split into $l_{nb}$ bands and the values from
each band are hashed again to create a number of keys. The original
record is added to a map associated with each of the keys. To perform
linkage, the algorithm iterates over the records in $\mathbf{D}_B$. Each
record is hashed as described above, to obtain a set of keys. Each key
is looked up in the data structure, and the associated records from
$\mathbf{D}_A$ added to the result set. Finally, the record from
$\mathbf{D}_B$ is compared in turn with each record in the result set,
with the pair being classified as a link or non-link based on their
distance.

@@

We now describe the datasets and method used in our
evaluation, as available from: \url{ADD\_URL\_TO\_GITHUB?}
%~\footnote{The raw data, additional figures and the source
%code required to run these experiments can be downloaded from: (link
%removed for double blind review).}.
\label{sec-data} We used three
datasets from two domains in our experiments, as summarised in
Table~\ref{table-datasets}. The first is \emph{Cora}~\cite{Cora2017},
which contains 1,295 records that refer to 112 machine learning
publications. Cora is commonly used as a benchmark dataset in the
literature for assessing linkage algorithms. Ground truth is provided
via a unique \emph{paper\_id} identifier of the form ``blum1993". In
this experiment linkage is performed over the same set of records (i.e.\
a de-duplication~\cite{Chr12}).

@@

@@

We perform linkage on the Cora dataset using all approaches presented in
this paper: brute force, traditional blocking, LSH and M-tree, using
several selected configurations for blocking and LSH. The distance
threshold is varied between 0 and 250~\footnote{Relatively high
Levenshtein edit distances are included since Cora includes a number of
low-similarity true matches.}. For traditional blocking, the following
attributes are used individually as blocking keys: \emph{author},
\emph{title}, \emph{venue}, \emph{location}, \emph{publisher} and
\emph{year}. We also use a combined blocking key comprising all
attributes.

\Cref{cora-quality} shows the precision, recall, and
F-measure~\cite{Chr12} for various thresholds. As expected, low
thresholds give high precision and low recall, and the reverse for high
thresholds. Brute force and M-tree give identical results, as expected.
The best linkage quality, with an F-measure of around 0.7, is achieved
by several linkers, including brute force, M-tree, blocking on
\emph{authors}, blocking on all attributes, and two of the LSH
configurations. All of these give similar overall results, apart from
blocking on \emph{authors}, which gives much better quality at very high
distance thresholds. This is due to the incomplete nature of the
approach, avoiding comparisons of significant numbers of high-similarity
non-matches and thus avoiding these becoming false positives and keeping
precision high.

@@

% Peter: Figure is too high - higher than page hight! ************
% So I changed the \vspace{5mm} to \vspace{4mm} and made the plots less % wide ************

\begin{figure}
\centering
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\vspace{-6mm}
\caption{Brute Force}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\vspace{-6mm}
\caption{M-tree}
\end{subfigure} \vspace{4mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\vspace{-6mm}
\caption{Blocking on `authors' attribute}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\vspace{-6mm}
\caption{Blocking on `title' attribute}
\end{subfigure} \vspace{4mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\vspace{-6mm}
\caption{Blocking on `publication year' attr.}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\vspace{-6mm}
\caption{Blocking using the union of all attrs.}
\end{subfigure} \vspace{4mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
\vspace{-6mm}
\caption{LSH using 2 bands of size 2}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
\vspace{-6mm}
\caption{LSH using 2 bands of size 10}
\end{subfigure} \vspace{4mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
\vspace{-6mm}
\caption{LSH using 10 bands of size 2}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
\vspace{-6mm}
\caption{LSH using 10 bands of size 10}
\end{subfigure}

\caption{Linkage results on the Cora dataset.}
\label{cora-quality}
\end{figure}

\begin{table}[t]
\caption{Linkage quality on Cora dataset with distance threshold
  $d= 70$.}
  \label{comparison-of-results-cora}
  \centering
  \begin{footnotesize}
  \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68  \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

@@

\Cref{demography-quality} plots (a) and (b) show the M-tree precision,
recall, and F-measure for various thresholds. In both datasets, the best
F-measure values are obtained with a low distance threshold of $d=2$.
Plots (c) and (d) compare the F-measure curves for M-tree with those
obtained from a range of LSH configurations. The best F-measure value
for M-tree is higher than that of any of the LSH configurations, for
both datasets. This demonstrates both the competitiveness of M-tree
with respect to linkage quality, and its important characteristic of
being parameter-free---the linkage quality is obtained without the
need to tune for the dataset.

@@

\begin{figure}[t]
\centering
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Kilmarnock dataset \label{kilmarnock-quality-mtree}}
\end{subfigure} \vspace{4mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\vspace{-6mm}
\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\vspace{-6mm}
\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
\end{subfigure}
\caption{Linkage results on the demographic datasets.}
\label{demography-quality}
\end{figure}

@@

\begin{table}[t]
\caption{Linkage quality on Skye dataset with distance threshold
$d= 2$.}
\label{comparison-of-results-demography-skye}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\

LSH-5-2  & 10,412,496                  & 0.64      & 0.11   & 0.19       \\

LSH-10-2 & 53,874,127                  & 0.68      & 0.36   & 0.47       \\

LSH-5-5  & 36,566                     & 0.76      & 0.01   & 0.01       \\

LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

\begin{table}[t]
\caption{Linkage quality on Kilmarnock dataset with distance
threshold $d= 2$.}
\label{comparison-of-results-demography-kili}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ 
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

We note that the good results yielded by both traditional blocking and
LSH are partly due to the fact that (in the limit) they tend towards
brute force as the number of records in the block increase. A second,
unexpected, result is that illustrated in
\Cref{comparison-of-results-cora}, namely that incomplete approaches
such as LSH can in some cases yield higher precision than that achieved
by a complete method such as M-tree. This is due to the incomplete
linker masking the inability of a classifier based solely on record
similarity to correctly classify high-similarity non-matches or
low-similarity matches.

@@

Different metrics can give different distributions of inter-record
distances, which can affect both the linkage results and the number of
comparisons made, and hence the efficiency of the algorithm. Datasets
with low variation in inter-record distances are said to have high
\textit{intrinsic dimensionality}, and tend to require high numbers of
comparisons. The Scottish vital event datasets, combined with
Levenshtein-based metrics, appear to have high intrinsic dimensionality.
We are now, therefore, investigating the application of various
different metrics to this domain, including Jensen-Shannon, Cosine and
Structured Entropic Distance, as described in~\cite{Connor2016}. We are
also investigating the applicability of a novel technique for
dimensionality reduction~\cite{Connor2017}.

>>>>

Record linkage is the process of identifying records that refer to the
same real-world entities in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair being classified as a link or
non-link depending on their similarity. Linkage is usually performed in
a three-step process: first, groups of similar candidate records are
identified using indexing, then pairs within the same group are compared
in more detail, and finally classified. Even state-of-the-art indexing
techniques, such as locality sensitive hashing, have potential
drawbacks. They may fail to group together some true matching records
with high similarity, or they may group records with low similarity,
leading to high computational overhead. We propose using \emph{metric
space indexing} (MSI) to perform \emph{complete} linkage, resulting in a
parameter-free process combining indexing, comparison and classification
into a single step delivering complete and efficient record linkage. An
evaluation on real-world data from several domains shows that linkage
using MSI can yield better quality than current indexing techniques,
with similar execution cost, without the need for domain knowledge or
trial and error to configure the process.

@@

To overcome data quality issues such as typographical errors and
variations (common in name and address values~\cite{Chr12}), approximate
string comparison functions (e.g. edit distance, the Jaro-Winkler
comparator, or Jaccard similarity~\cite{Chr12}) are used to compare
record pairs, leading to a vector of similarities (one similarity per
attribute compared) for each pair. These are used to classify the record
pairs into \emph{links} (where it is assumed both records correspond to
the same real-world entity) and \emph{non-links} (where they are assumed
to correspond to different entities). Various classification methods
have been employed in record linkage~\cite{Chr12,Don15}, ranging from
simple threshold-based to sophisticated clustering, supervised
classification, and active learning approaches~\cite{Wan15}.

@@

Traditional linkage systems that perform indexing prior to comparison
and classification (on the left in Fig.~\ref{fig-rl-process}) add a
further complexity. Indexing, comparison and classification are often
conducted using algorithms and parameters selected using domain
expertise, followed by manual assessment of the linkage
outcomes~\cite{Chr12}. If the resulting link quality is too low for a
certain application, the process is repeated with different parameter
settings or algorithms, giving a time-consuming iterative
process~\cite{Fis15}. The choice of an appropriate indexing technique
as well as suitable parameter settings (including which attributes to
use in indexing) will significantly affect the final linkage outcome.

@@

Metric space indexing (MSI) is a complete technique with lower
computational cost than a brute force approach. It allows indexing,
comparison and classification to be combined into a single step (on the
right in Fig.~\ref{fig-rl-process}), making the process simpler, more
efficient and more effective than incomplete approaches.

The motivation for this work is the Digitising Scotland
project~\cite{Dibben2012}, which aims to transcribe and link all civil
registration events recorded in Scotland between 1856 and 1973. This
dataset will include around 14 million birth records, 11 million death
records and 4 million marriage records.

@@

Locality sensitive hashing (LSH), proposed for efficient
nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
been used for record linkage indexing. Attribute values are hashed
multiple times, and blocks are created from those records that share
some hash values. \emph{HARRA}~\cite{Kim10} is a linkage approach based
on MinHash~\cite{Broder1997} and LSH which blocks, compares, and then
merges linked records iteratively. \cite{Steorts2014} evaluates two LSH
variations, concluding that to get good results, they must be tuned to
the datasets. This requires good ground truth data which may be
unavailable in real-world applications or expensive to obtain.

@@

\textbf{Brute force:} Every record in $\mathbf{D}_A$ is compared with
every record in $\mathbf{D}_B$. Each pair is classified as a link if
the distance between the records is less than or equal to the
threshold $d$. This always finds all links, with complexity
$ O(|\mathbf{D}_A| \cdot |\mathbf{D}_B|)$.

\textbf{Traditional blocking:} The parameters are the set of blocking
keys and (optionally) the phonetic encodings applied to each attribute.
These are selected as described in~\cite{Chr12b}, exploiting knowledge
of the domain and of the data, and chosen with the intention of giving
the best possible results. Each record in $\mathbf{D}_A$ is placed into
the appropriate block based on its blocking key value. The algorithm
then iterates over the records in $\mathbf{D}_B$, and for each one
compares it with each of the records from $\mathbf{D}_A$ in the block
with the same blocking key value.

\textbf{LSH-MinHash:} The parameters for LSH-Minhash
are~\cite{Broder1997} \emph{shingle size} ($l_{ss}$), \emph{band size}
($l_{bs}$) and \emph{number of bands} ($l_{nb}$). First, the attributes
of each record in $\mathbf{D}_A$ are concatenated, and the result
\emph{shingled} into a set of n-grams with $n = l_{ss}$. Next, a set of
deterministically generated hash functions is applied to each n-gram in
the set and the smallest result (the MinHash) of each hash application
is added to a signature for the record. The number of hashes used, and
thus the size of the signature, is set to $l_{nb} \times l_{bs}$.
Finally, the signature is split into $l_{nb}$ bands and the values from
each band are hashed again to create a number of keys. The original
record is added to a map associated with each of the keys. To perform
linkage, the algorithm iterates over the records in $\mathbf{D}_B$. Each
record is hashed as described above, to obtain a set of keys. Each key
is looked up in the data structure, and the associated records from
$\mathbf{D}_A$ added to the result set. Finally, the record from
$\mathbf{D}_B$ is compared in turn with each record in the result set,
with the pair being classified as a link or non-link based on their
distance.

@@

We now describe the datasets and method used in our
evaluation~\footnote{The raw experimental data, additional figures and
the source code required to run these experiments can be downloaded
from the following repository. \\
\url{http://github.com/digitisingscotland/pakdd2018-metric-linkage}.}.
We used three datasets from two
domains in our experiments, as summarised in Table~\ref{table-datasets}.
The first is \emph{Cora}~\cite{Cora2017}, which contains 1,295 records
that refer to 112 machine learning publications. Cora is commonly used
as a benchmark dataset in the literature for assessing linkage
algorithms. Ground truth is provided via a unique \emph{paper\_id}
identifier of the form ``blum1993". In this experiment linkage is
performed over the same set of records (i.e.\ a
de-duplication~\cite{Chr12}).

@@

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

@@

\begin{figure}
\centering
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\vspace{-6mm}
\caption{Brute Force}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\vspace{-6mm}
\caption{M-tree}
\end{subfigure} \vspace{3mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\vspace{-6mm}
\caption{Blocking on `authors' attribute}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\vspace{-6mm}
\caption{Blocking on `title' attribute}
\end{subfigure} \vspace{3mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\vspace{-6mm}
\caption{Blocking on `publication year' attr.}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\vspace{-6mm}
\caption{Blocking using the union of all attrs.}
\end{subfigure} \vspace{3mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
\vspace{-6mm}
\caption{LSH using 2 bands of size 2}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
\vspace{-6mm}
\caption{LSH using 2 bands of size 10}
\end{subfigure} \vspace{3mm}

\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
\vspace{-6mm}
\caption{LSH using 10 bands of size 2}
\end{subfigure}%
~~
\begin{subfigure}{.47\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
\vspace{-6mm}
\caption{LSH using 10 bands of size 10}
\end{subfigure}
\caption{Linkage results on the Cora dataset.}
\label{cora-quality}
\end{figure}

We perform linkage on the Cora dataset using all approaches presented in
this paper: brute force, traditional blocking, LSH and M-tree, using
several selected configurations for blocking and LSH. The distance
threshold is varied between 0 and 250~\footnote{Relatively high
Levenshtein edit distances are included since Cora contains a number of
low-similarity true matches.}. For traditional blocking, the following
attributes are used individually as blocking keys: \emph{author},
\emph{title}, \emph{venue}, \emph{location}, \emph{publisher} and
\emph{year}. We also use a combined blocking key comprising all
attributes.

Figure~\ref{cora-quality} shows the precision, recall, and
F-measure~\cite{Chr12} for various thresholds~\footnote{Noting that
recent research identifies some problematic aspects with using the
F-measure to compare record linkage procedures at different similarity
thresholds~\cite{Han18}.}. As expected, low thresholds give high
precision and low recall, and the reverse for high thresholds. Brute
force and M-tree give identical results, as expected. The best linkage
quality, with an F-measure of around 0.7, is achieved by several
linkers, including brute force, M-tree, blocking on \emph{authors},
blocking on all attributes, and two of the LSH configurations. All of
these give similar overall results, apart from blocking on
\emph{authors}, which gives much better quality at very high distance
thresholds. This is due to the incomplete nature of the approach,
avoiding comparisons of significant numbers of high-similarity
non-matches and thus avoiding these becoming false positives and keeping
precision high.

@@

\begin{table}[t]
\caption{Linkage quality on Cora dataset with distance threshold $d= 70$.}
\label{comparison-of-results-cora}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & ~~~~~~~~~~~Recall & ~~~~~~F-measure \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Brute Force     & 1,677,025 & 0.84 & 0.57 & 0.68 \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree          &   902,693 & 0.84 & 0.57 & 0.68 \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2         &   192,199 & 0.95 & 0.47 & 0.63 \\
LSH-5-2         &   342,849 & 0.91 & 0.55 & 0.69 \\
LSH-10-2        &   513,947 & 0.88 & 0.57 & 0.69 \\
LSH-2-5         &    14,329 & 0.99 & 0.28 & 0.43 \\
LSH-5-5         &    22,057 & 0.99 & 0.36 & 0.53 \\
LSH-10-5        &    26,167 & 0.98 & 0.40 & 0.57 \\
LSH-2-10        &     4,711 &    1 & 0.15 & 0.27 \\
LSH-5-10        &     6,501 &    1 & 0.19 & 0.32 \\
LSH-10-10       &    10,627 & 0.99 & 0.27 & 0.43 \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Block-year      &   115,893 & 0.99 & 0.35 & 0.51 \\
Block-authors   &    11,039 & 0.94 & 0.16 & 0.28 \\
Block-title     &    27,407 & 0.95 & 0.42 & 0.58 \\
Block-venue     &    36,647 & 0.85 & 0.29 & 0.44 \\
Block-location  & 1,009,957 & 0.83 & 0.43 & 0.57 \\
Block-publisher &   833,079 & 0.85 & 0.44 & 0.58 \\
Block-combined  & 1,214,269 & 0.84 & 0.56 & 0.67 \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

\begin{figure}[t]
\centering
\begin{subfigure}{.46\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Isle of Skye \\ dataset\label{skye-quality-mtree}}
\end{subfigure}%
~~
\begin{subfigure}{.46\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Kilmarnock \\ dataset \label{kilmarnock-quality-mtree}}
\end{subfigure} \vspace{1mm}

\begin{subfigure}{.46\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\vspace{-6mm}
\caption{F-measure on Isle of Skye dataset \\ with M-tree and all
LSH configurations}
\end{subfigure}%
~~
\begin{subfigure}{.46\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\vspace{-6mm}
\caption{F-measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
\end{subfigure}
\caption{Linkage results on the demographic datasets.}
\label{demography-quality}
\end{figure}

@@

Figure~\ref{demography-quality} plots (a) and (b) show the M-tree precision,
recall, and F-measure for various thresholds. In both datasets, the best
F-measure values are obtained with a low distance threshold of $d=2$.
Plots (c) and (d) compare the F-measure curves for M-tree with those
obtained from a range of LSH configurations. The best F-measure value
for M-tree is higher than that of any of the LSH configurations, for
both datasets. This demonstrates both the competitiveness of M-tree
with respect to linkage quality, and its important characteristic of
being parameter-free---the linkage quality is obtained without the
need to tune for the dataset.

@@

\begin{table}[t]
\caption{Linkage quality on Isle of Skye dataset with distance threshold $d=2$.}
\label{comparison-of-results-demography-skye}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & ~~~~~~~~~~~Recall & ~~~~~~F-measure \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree   & 102,318,525 & 0.65 & 0.46 & 0.54 \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  &   3,109,250 & 0.63 & 0.03 & 0.06 \\
LSH-5-2  &  10,412,496 & 0.64 & 0.11 & 0.19 \\
LSH-10-2 &  53,874,127 & 0.68 & 0.36 & 0.47 \\
LSH-5-5  &      36,566 & 0.76 & 0.01 & 0.01 \\
LSH-10-5 &     129,873 & 0.72 & 0.01 & 0.02 \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

\begin{table}[t]
\caption{Linkage quality on Kilmarnock dataset with distance threshold $d=2$.}
\label{comparison-of-results-demography-kili}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & ~~~~~~~~~~~Recall & ~~~~~~F-measure \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree   & 514,871,153 & 0.76 & 0.45 & 0.57 \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  &  99,145,887 & 0.81 & 0.16 & 0.27 \\
LSH-5-2  & 130,721,338 & 0.79 & 0.23 & 0.36 \\
LSH-10-2 & 177,168,848 & 0.79 & 0.36 & 0.49 \\
LSH-5-5  &     239,368 & 0.84 & 0.01 & 0.02 \\
LSH-10-5 &     855,431 & 0.87 & 0.02 & 0.03 \\ 
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

We note that the good results obtained by both traditional blocking and
LSH are partly due to the fact that (in the limit) they tend towards
brute force as the number of records in the blocks increase. A second,
unexpected, result is that illustrated in
Table~\ref{comparison-of-results-cora}, namely that incomplete approaches
such as LSH can in some cases yield higher precision than that achieved
by a complete method such as M-tree. This is due to the incomplete
linker masking the inability of a classifier based solely on record
similarity to correctly classify high-similarity non-matches or
low-similarity matches.

@@

Different metrics can give different distributions of inter-record
distances, which can affect both the linkage results and the number of
comparisons made, and hence the efficiency of the algorithm. Datasets
with low variation in inter-record distances are said to have high
\textit{intrinsic dimensionality}, and tend to require high numbers of
comparisons. The Scottish vital event datasets~\cite{reid2002}, combined
with Levenshtein-based metrics, appear to have high intrinsic
dimensionality. We are now, therefore, investigating the application of
various different metrics to this domain, including Jensen-Shannon,
Cosine and Structured Entropic Distance, as described
in~\cite{Connor2016}. We are also investigating the applicability of a
novel technique for dimensionality reduction~\cite{Connor2017}.
